---
title: "HLS"
author: "Apurv Srivastav"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download R
```{r}
data<-read.csv("hls.csv")
```

```{r}
str(data)
```

```{r}
### remove restrivted
datasub<- data[sapply(data, function(x) x[1] != -5)]
colnames(datasub)
```

```{r}
### remove f2 and f3 variables (second and third followups)
datasub1<-datasub[,-c(4,8:11,212:616,1520:1939)]
colnames(datasub1)
```


### Identification Variables

STU_ID (1), STRAT_ID (2), PSU (3), F1UNIV1 (4), F1UNIV2A (5), F1UNIV2B (6), G10COHRT(7),G12COHRT(8), BYSTUWT(9)

### Variables to Keep

BYSEX (10), BYRACE (11), BYDOB_P (15), BYPARACE (16), BYGNSTAT(22), BYMOTHED(24), BYPATHED (25), BYGPARED (26), BYOCCUM (27), BYOCCUF (28), BYNONUSG (36), BYHOMLIT (37), BYRISKFC (38), BYSTEXP (39), BYPARASP(40), BYSCHPRG(43), BYTXSTAT(46), BYTXACC (53), $\bf BYTXMSTD\ (62)$, $\bf BYTXMSTD\ (70)$, BYCONEXP (77), BYINSTMO(78), BYACTCTL(79), BYSTPREP (80), BYWRTNGA (81), BYNSPORT (96), BYXTRACU (97), BYHMWRK (98), BYTVVIGM (99), BYERACE (102), BYTEHDEG (103), BYMRACE (104), BYTMHDEG (105), BYSCTRL (110), BYURBAN (111), BYREGION (112), BY10FLP (116), F1EVERDO (165), $\bf F1TXMSTD (172)$, F1HIMATH (182)

### Variables to keep 1 after analysis analysis

BYSTLANG (12), BYHOMLNG (13),  or BYSTLNG2 (14)

 BYPARLNG (17) or BYPLANG (18)

BYFCOMP (19), BYSIBSTR (20), or BYSIBHOM (21)

BYSES1 (30) or BYSES2 (32)

BYGRDRPT (34) or BYLGRRPT (35)

BYWORKSY (100) or BYWRKHRS (101)

BYS33A (313) and BYS33B (314) (combine)

### Variables to Delete

BYPARED (23), BYSES1QU (31), BYSES2QU (33), BYOCCHS (41), BYOCC30 (42), BYSQSTAT (44), BYQXDATP (45), BYTEQFLG(47), BYPQSTAT (48), BYTXPAFG (49), BYADMFLG(50), BYLMCFLG (51), BYIEPFLG (52), BYTXCSTD (54), BYTXCQU (55), BYNELS2M (56), BYNELS2R (57), BYNELS0M (58), BYPISAME (59), BYPISARE (60), BYTXMIRR (61), BYTXMQU (63), BYTX1MPP (64), BYTX2MPP (65), BYTX3MPP (66), BYTX4MPP (67), BYTX5MPP (68), BYTXRIRR (69), BYTXRQU (71), BYTX1RPP (72), BYTX2RPP (73), BYTX3RPP (74), BYMATHSE (75), BYENGLSE (76), BYSF1RCE (82), BYSF2RCE (83), BYSF3RCE (84), BYFRRACE (85), BYFRGRIM (86), BYBASEBL (87), BYSOFTBL (88), BYBSKTBL (89), BYFOOTBL (90), BYSOCCER (91), BYTEAMSP (92), BYSOLOSP (93), BYCHRDRL (94), BYTEAQUA (106), BYTSTREL (107), BYG10EP (108), BYSCENP (109), BYREGURB (113), BYREGCTL(114), BYSPANP (115), BYSCSAF1 (117), BYSCSAF2(118), BYACCLIM (119), BYSEXIM (120), BYSTLNIM (121), BYFCMPIM (122), BYMOEDIM (123), BYFAEDIM (124), BYOCCMIM (125), BYOCCFIM (126), BYINCMIM (127), BYSTEXIM (128), BYPASPIM (129), BYSCHPIM (130), BYTESTIM (131), BYMATHIM (132), BYREADIM (133) F1QWT (134), F1PNLWT (135), F1SEX (136), F1RACE (137), F1STLANG (138), F1HOMLNG (139), F1DOB_P  (140), F1FCOMP (141), F1PARED (142),  F1MOTHED (143), F1FATHED (144), F1OCCUM (145), F1OCCUF  (146), F1SES1 (147), F1SES1QU (148), F1SES2 (149), F1SES2QU (150), F1STEXP (151),F1BYEDEX (152), F1OCCHS (153), F1OCC30 (154), F1QSTAT (155), F1TSTAT (156), F1PNLFLG (157), F1BYTFLG (158), F1QMODE (159), F1STQFLG (160), F1TRQFLG (161), F1HOQFLG (162), F1EGQFLG (163), F1DOQFLG (164), F1DOSTAT (166),  F1TXACC (167), F1CTLPTN (168), F1GRADE (169), F1TXMBIR (170), F1TXM1IR (171), F1TXMQU (173), F1NELS2M (174), F1TX1MPP (175), F1TX2MPP (176), F1TX3MPP (177), F1TX4MPP (178), F1TX5MPP (179), F1BYMTGN (180), F1QXDATP (181), Variables 183- 312, 315-2254        
###### 

```{r}
### eliminating variables
datasub2<-datasub1[,-c(31,33,41,42,44,45,47,48:52,54:61,63:69,71:76,82:94, 106:109,113:115,117:164, 166:171,173:181, 183:312,315:1394, 1396:1481, 1483:2254)]
str(datasub2)
```



```{r}
datasub3<- datasub2[,-c(2,3,4:6,8,9, 30:31,36,40,44,49, 63:66) ]
str(datasub3)
```

```{r,message=FALSE,warning=FALSE}
library(dplyr)
datasub4<- mutate(datasub3, BYAPIB = ifelse(as.factor(BYS33A)==c("-9","-8","-7","-6","-4")|as.factor(BYS33B)==c("-9","-8","-7","-6","-4"),"-9",ifelse(as.factor(BYS33A)=="1"|as.factor(BYS33B)=="1", "1","0")))
datasub4$BYAPIB<-as.factor(datasub4$BYAPIB)
summary(datasub4$BYAPIB)
```

```{r,message=FALSE,warning=FALSE}
library(plyr)
datasub4$BYERACE<-revalue(as.factor(datasub4$BYERACE), c("-9"="-8"))
summary(datasub4$BYERACE)
datasub5<- mutate(datasub4, BYERCEMTCH = ifelse(as.factor(BYRACE)==c("-8","-4")|BYERACE==c("-8","-4"),"-9",ifelse(as.factor(BYRACE)==BYERACE, "1","0")))
datasub5$BYERCEMTCH<-as.factor(datasub5$BYERCEMTCH)
summary(datasub5$BYERCEMTCH)
```



```{r, message=FALSE, warning=FALSE}
library(plyr)
datasub5$BYMRACE<-revalue(as.factor(datasub5$BYMRACE), c("-9"="-8"))
summary(datasub5$BYMRACE)
datasub6<- mutate(datasub5, BYMRCEMTCH = ifelse(as.factor(BYRACE)==c("-8","-4")|BYMRACE==c("-8","-4"),"-9",ifelse(as.factor(BYRACE)==BYMRACE, "1","0")))
datasub6$BYMRCEMTCH<-as.factor(datasub6$BYMRCEMTCH)
summary(datasub6$BYMRCEMTCH)
```

```{r}
library(plyr)
datasub6$BYTE22<-revalue(as.factor(datasub6$BYTE22), c("-9"="-8"))
summary(datasub6$BYTE22)
datasub7<- mutate(datasub6, BYESEXMTCH = ifelse(as.factor(BYSEX)==c("-8","-4")|BYTE22==c("-8","-4"),"-9",ifelse(as.factor(BYSEX)==as.factor(BYTE22), "1","0")))
datasub7$BYESEXMTCH<-as.factor(datasub7$BYESEXMTCH)
summary(datasub7$BYESEXMTCH)
```

```{r}
library(plyr)
datasub7$BYTM22<-revalue(as.factor(datasub7$BYTM22), c("-9"="-8"))
summary(datasub7$BYTM22)
datasub8<- mutate(datasub7, BYMSEXMTCH = ifelse(as.factor(BYSEX)==c("-8","-4")|BYTM22==c("-8","-4"),"-9",ifelse(as.factor(BYSEX)==as.factor(BYTM22), "1","0")))
datasub8$BYMSEXMTCH<-as.factor(datasub8$BYMSEXMTCH)
summary(datasub8$BYMSEXMTCH)
```




```{r}
colnames(datasub8)
```


```{r}
datasub9<-datasub8[,-c(3,4,43,45,50,51,52,53)]
colnames(datasub9)
test<-datasub9[,-c(2,7,15:17)]
```

```{r}
test$BYESEXMTCH<- replace(test$BYESEXMTCH, which(test$BYESEXMTCH=="-9"),NA)
test$BYMSEXMTCH<- replace(test$BYMSEXMTCH, which(test$BYMSEXMTCH=="-9"),NA)
test$BYERCEMTCH<- replace(test$BYERCEMTCH, which(test$BYERCEMTCH=="-9"),NA)
test$BYMRCEMTCH<- replace(test$BYMRCEMTCH, which(test$BYMRCEMTCH=="-9"),NA)
test$BYSTLANG<- replace(test$BYSTLANG, which(test$BYSTLANG<0),NA)
test$BYHOMLNG<- replace(test$BYHOMLNG, which(test$BYHOMLNG<0),NA)
test$BYSTLNG2<- replace(test$BYSTLNG2, which(test$BYSTLNG2<0),NA)
test$BYDOB_P<- replace(test$BYDOB_P, which(test$BYDOB_P<0),NA)
test$BYPARLNG<- replace(test$BYPARLNG, which(test$BYPARLNG<0),NA)
test$BYPLANG<- replace(test$BYPLANG, which(test$BYPLANG<0),NA)
test$BYFCOMP<- replace(test$BYFCOMP, which(test$BYFCOMP<0),NA)
test$BYSIBSTR<- replace(test$BYSIBSTR, which(test$BYSIBSTR<0),NA)
test$BYSIBHOM<- replace(test$BYSIBHOM, which(test$BYSIBHOM<0),NA)
test$BYGNSTAT<- replace(test$BYGNSTAT, which(test$BYGNSTAT<0),NA)


test$BYPARED<- replace(test$BYPARED, which(test$BYPARED<0),NA)
test$BYOCCUM<- replace(test$BYOCCUM, which(test$BYOCCUM<0),NA)
test$BYOCCUF<- replace(test$BYOCCUF, which(test$BYOCCUF<0),NA)
test$BYINCOME<- replace(test$BYINCOME, which(test$BYINCOME<0),NA)
test$BYGRDRPT<- replace(test$BYGRDRPT, which(test$BYGRDRPT<0),NA)
test$BYLGRRPT<- replace(test$BYLGRRPT, which(test$BYLGRRPT<0),NA)
test$BYLGRRPT<- replace(test$BYLGRRPT, which(test$BYLGRRPT>97),NA)
test$BYNONUSG<- replace(test$BYNONUSG, which(test$BYNONUSG<0),NA)
test$BYSTEXP<- replace(test$BYSTEXP, which(test$BYSTEXP<0),NA)
test$BYPARASP<- replace(test$BYPARASP, which(test$BYPARASP<0),NA)
test$BYSCHPRG<- replace(test$BYSCHPRG, which(test$BYSCHPRG<0),NA)


test$BYTXACC<- replace(test$BYTXACC, which(test$BYTXACC<0),NA)
test$BYTXMSTD<- replace(test$BYTXMSTD, which(test$BYTXMSTD<0),NA)
test$BYTXRSTD<- replace(test$BYTXRSTD, which(test$BYTXRSTD<0),NA)
test$BYINSTMO<- replace(test$BYINSTMO, which(test$BYINSTMO<=-4),NA)
test$BYACTCTL<- replace(test$BYACTCTL, which(test$BYACTCTL<=-4),NA)
test$BYSTPREP<- replace(test$BYSTPREP, which(test$BYSTPREP<=-4),NA)
test$BYWRTNGA<- replace(test$BYWRTNGA, which(test$BYWRTNGA<=-4),NA)
test$BYNSPORT<- replace(test$BYNSPORT, which(test$BYNSPORT<0),NA)
test$BYXTRACU<- replace(test$BYXTRACU, which(test$BYXTRACU<0),NA)
test$BYHMWRK<- replace(test$BYHMWRK, which(test$BYHMWRK<0),NA)
test$BYWORKSY<- replace(test$BYWORKSY, which(test$BYWORKSY<0),NA)


test$BYWRKHRS<- replace(test$BYWRKHRS, which(test$BYWRKHRS<0),NA)
test$BYTEHDEG<- replace(test$BYTEHDEG, which(test$BYTEHDEG<0),NA)
test$BYTMHDEG<- replace(test$BYTMHDEG, which(test$BYTMHDEG<0),NA)
test$BYSCTRL<- replace(test$BYSCTRL, which(test$BYSCTRL<0),NA)
test$BYREGION<- replace(test$BYREGION, which(test$BYREGION<0),NA)
test$BYAPIB<- replace(test$BYAPIB, which(test$BYAPIB=="-9"),NA)
test$BYERCEMTCH<- replace(test$BYERCEMTCH, which(test$BYERCEMTCH=="-9"),NA)
test$BYMRCEMTCH<- replace(test$BYMRCEMTCH, which(test$BYMRCEMTCH=="-9"),NA)
test$BYTVVIGM<- replace(test$BYTVVIGM, which(test$BYTVVIGM<0),NA)
test$BYTVVIGM<- replace(test$BYTVVIGM, which(test$BYTVVIGM>=99),NA)
test$BYURBAN<- replace(test$BYURBAN, which(test$BYURBAN<0),NA)
test$BYGRDRPT<- replace(test$BYGRDRPT, which(test$BYGRDRPT>97),NA)
test$BYHOMLIT<- replace(test$BYHOMLIT, which(test$BYHOMLIT<0),NA)

```

```{r}
test$BYERCEMTCH<-droplevels(test$BYERCEMTCH)
test$BYAPIB<-droplevels(test$BYAPIB)
test$BYMRCEMTCH<-droplevels(test$BYMRCEMTCH)
test$BYESEXMTCH<-droplevels(test$BYESEXMTCH)
summary(test$BYESEXMATCH)
test$BYMSEXMTCH<-droplevels(test$BYMSEXMTCH)
summary(test$BYMSEXMTCH)

```




```{r}
colnames(test)
```


```{r}
test[,c(1:23,30:45)]<-lapply(test[,c(1:23,30:45)],factor)
str(test)
summary(test)

```



```{r}
attach(test)
```

```{r, warning=FALSE, message=FALSE}
## English score results
summary(aov(BYTXRSTD~BYESEXMTCH))
summary(aov(BYTXRSTD~BYERCEMTCH))
summary(aov(BYTXRSTD~BYDOB_P))
summary(aov(BYTXRSTD~BYFCOMP))
summary(aov(BYTXRSTD~BYPARED))
summary(aov(BYTXRSTD~BYOCCUM))
summary(aov(BYTXRSTD~BYOCCUF))
summary(aov(BYTXRSTD~BYINCOME))
summary(aov(BYTXRSTD~BYNONUSG))
summary(aov(BYTXRSTD~BYHOMLIT))
summary(aov(BYTXRSTD~BYSTEXP))
summary(aov(BYTXRSTD~BYPARASP))
summary(aov(BYTXRSTD~BYSCHPRG))
summary(aov(BYTXRSTD~BYTXACC))
summary(aov(BYTXRSTD~BYNSPORT))
summary(aov(BYTXRSTD~BYXTRACU))
summary(aov(BYTXRSTD~BYHMWRK))
summary(aov(BYTXRSTD~BYTVVIGM))
summary(aov(BYTXRSTD~BYTEHDEG))
summary(aov(BYTXRSTD~BYSCTRL))
summary(aov(BYTXRSTD~BYURBAN))
summary(aov(BYTXRSTD~BYREGION))
summary(aov(BYTXRSTD~BYAPIB))
```

```{r}
summary(aov(BYTXRSTD~BYSTLANG))
summary(aov(BYTXRSTD~BYHOMLNG))
summary(aov(BYTXRSTD~BYSTLNG2))
```

```{r}
summary(aov(BYTXRSTD~BYPARLNG))
summary(aov(BYTXRSTD~BYPLANG))
```

```{r}
summary(aov(BYTXRSTD~BYSIBSTR))
summary(aov(BYTXRSTD~BYSIBHOM))
```

```{r}
summary(aov(BYTXRSTD~BYGRDRPT))
summary(aov(BYTXRSTD~BYLGRRPT))
```

```{r}
summary(aov(BYTXRSTD~BYWORKSY))
summary(aov(BYTXRSTD~BYWRKHRS))
```


```{r, warning=FALSE, message=FALSE}
## MATH score results
summary(aov(BYTXMSTD~BYMSEXMTCH))
summary(aov(BYTXMSTD~BYMRCEMTCH))
summary(aov(BYTXMSTD~BYDOB_P))
summary(aov(BYTXMSTD~BYFCOMP))
summary(aov(BYTXMSTD~BYPARED))
summary(aov(BYTXMSTD~BYOCCUM))
summary(aov(BYTXMSTD~BYOCCUF))
summary(aov(BYTXMSTD~BYINCOME))
summary(aov(BYTXMSTD~BYNONUSG))
summary(aov(BYTXMSTD~BYHOMLIT))
summary(aov(BYTXMSTD~BYSTEXP))
summary(aov(BYTXMSTD~BYPARASP))
summary(aov(BYTXMSTD~BYSCHPRG))
summary(aov(BYTXMSTD~BYTXACC))
summary(aov(BYTXMSTD~BYNSPORT))
summary(aov(BYTXMSTD~BYXTRACU))
summary(aov(BYTXMSTD~BYHMWRK))
summary(aov(BYTXMSTD~BYTVVIGM))
summary(aov(BYTXMSTD~BYTMHDEG))
summary(aov(BYTXMSTD~BYSCTRL))
summary(aov(BYTXMSTD~BYURBAN))
summary(aov(BYTXMSTD~BYREGION))
summary(aov(BYTXMSTD~BYAPIB))

```

```{r}
summary(aov(BYTXMSTD~BYSTLANG))
summary(aov(BYTXMSTD~BYHOMLNG))
summary(aov(BYTXMSTD~BYSTLNG2))
```

```{r}
summary(aov(BYTXMSTD~BYPARLNG))
summary(aov(BYTXMSTD~BYPLANG))
```

```{r}
summary(aov(BYTXMSTD~BYSIBSTR))
summary(aov(BYTXMSTD~BYSIBHOM))
```

```{r}
summary(aov(BYTXMSTD~BYGRDRPT))
summary(aov(BYTXMSTD~BYLGRRPT))
```

```{r}
summary(aov(BYTXMSTD~BYWORKSY))
summary(aov(BYTXMSTD~BYWRKHRS))
```

```{r}
cor.test(BYACTCTL, BYTXMSTD,  method = "pearson")
cor.test(BYACTCTL, BYTXRSTD,  method = "pearson")
```

```{r}
cor.test(BYINSTMO, BYTXMSTD,  method = "pearson")
cor.test(BYINSTMO, BYTXRSTD,  method = "pearson")
```

```{r}
cor.test(BYSTPREP, BYTXMSTD,  method = "pearson")
cor.test(BYSTPREP, BYTXRSTD,  method = "pearson")
```

```{r}
cor.test(BYWRTNGA, BYTXMSTD,  method = "pearson")
cor.test(BYWRTNGA, BYTXRSTD,  method = "pearson")
```

## BYWORKSY, BYLGRRPT

```{r}
colnames(test)
datasub10<-test[,-c(5,17,34)]
colnames(datasub10)
attach(datasub10)
```

## chi sqaure separated by type of variable

```{r}
library(car)
#langmodm<-lm(BYTXMSTD~BYSTLANG+BYHOMLNG+BYSTLNG2+BYPARLNG+BYPLANG)
#langmode<-lm(BYTXRSTD~BYSTLANG+BYHOMLNG+BYSTLNG2+BYPARLNG+BYPLANG)

### Choose English Fluency (BYSTLNG2)

chisq.test(table(BYSTLANG,BYHOMLNG))
chisq.test(table(BYSTLANG,BYSTLNG2))
chisq.test(table(BYSTLANG,BYPARLNG))
chisq.test(table(BYSTLANG,BYPLANG))
chisq.test(table(BYSTLNG2,BYHOMLNG))
chisq.test(table(BYSTLNG2,BYPARLNG))
chisq.test(table(BYSTLNG2,BYPLANG))
chisq.test(table(BYHOMLNG,BYPARLNG))
chisq.test(table(BYHOMLNG,BYPLANG))
chisq.test(table(BYPARLNG,BYPLANG))
```

```{r}
# Family STructure Chi square

### Keep family structure (fcomp) and sibling number (BYSIBHOM)

chisq.test(table(BYFCOMP,BYSIBSTR))
chisq.test(table(BYFCOMP,BYSIBHOM))
chisq.test(table(BYFCOMP,BYGNSTAT))
chisq.test(table(BYSIBSTR,BYSIBHOM))
chisq.test(table(BYSIBSTR,BYGNSTAT))
chisq.test(table(BYSIBHOM,BYGNSTAT))
```

```{r}
### Parental stautus (education and job)

### BYPARED and BYINCOME kept (parents highest degree and income)


chisq.test(table(BYPARED,BYOCCUM))
chisq.test(table(BYPARED,BYINCOME))
chisq.test(table(BYOCCUM,BYINCOME))
chisq.test(table(BYPARED,BYOCCUF))
chisq.test(table(BYOCCUF,BYINCOME))



```

```{r}
### Education Variables

### Keep BYGRDRPT, BYNONUSG, BYSCHPRG

chisq.test(table(BYGRDRPT,BYNONUSG))
chisq.test(table(BYGRDRPT,BYHOMLIT))
chisq.test(table(BYGRDRPT,BYSCHPRG))
chisq.test(table(BYNONUSG,BYHOMLIT))
chisq.test(table(BYNONUSG,BYSCHPRG))
chisq.test(table(BYHOMLIT,BYSCHPRG))
```

```{r}

### keep BYACTCTL, BYSTPREP, BYWRTNGA (self rated effort, class preparation, writing level by teacher)

cor.test(BYINSTMO, BYACTCTL,  method = "pearson")
cor.test(BYINSTMO, BYSTPREP,  method = "pearson")
cor.test(BYINSTMO, BYWRTNGA,  method = "pearson")
cor.test(BYACTCTL, BYSTPREP,  method = "pearson")
cor.test(BYACTCTL, BYWRTNGA,  method = "pearson")
cor.test(BYSTPREP, BYWRTNGA,  method = "pearson")
```

```{r}
### Time spent

#### Keep BYHMWRK and BYWRKHRS (Homework hours and WOrk hours)

chisq.test(table(BYNSPORT,BYXTRACU))
chisq.test(table(BYNSPORT,BYHMWRK))
chisq.test(table(BYNSPORT,BYTVVIGM))
chisq.test(table(BYNSPORT,BYWRKHRS))
chisq.test(table(BYXTRACU,BYHMWRK))
chisq.test(table(BYXTRACU,BYTVVIGM))
chisq.test(table(BYXTRACU,BYWRKHRS))
chisq.test(table(BYHMWRK,BYTVVIGM))
chisq.test(table(BYHMWRK,BYWRKHRS))
chisq.test(table(BYTVVIGM,BYWRKHRS))
```

```{r}
### Type of School
chisq.test(table(BYSCTRL,BYURBAN))
chisq.test(table(BYSCTRL,BYREGION))
chisq.test(table(BYURBAN,BYREGION))
```

### Final Variables
```{r}
colnames(datasub10)
fdata<-datasub10[, c(1,2,7, 9, 11, 14, 15, 16, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 30, 32,33,34,35,36,37,38,39,40,41,42)]
str(fdata)

```
















```{r}
library(car)
fdata$BYHMWRK<-as.numeric(fdata$BYHMWRK)
```

### Redefine variables
```{r}
set.seed(22)
library(car)
### FCOMP now mother and father vs any other family structure
fdata$BYFCOMP<-recode(fdata$BYFCOMP, "c('2','3','4','5','6','7','8','9')='2'")
summary(fdata$BYFCOMP)

### BYSIMHOM is 0 or 1 sibling vs 2+ siblings
fdata$BYSIBHOM<-recode(fdata$BYSIBHOM, "c('0','1')='1';c('2','3','4','5','6','7')='2'")
summary(fdata$BYSIBHOM)

### BYPARED is High school or less, college dropout or 2 year program, 4 year college degree and igher
fdata$BYPARED<-recode(fdata$BYPARED, "c('1','2')='0';c('3','4','5')='1';c('6','7','8')='2'")
summary(fdata$BYPARED)

### BYINCOME is now 35K or less, 35K-75K, 75K or more
fdata$BYINCOME<-recode(fdata$BYINCOME, "c('1','2','3','4','5','6','7','8')='0';c('9','10')='1';c('11','12','13')='2'")
summary(fdata$BYINCOME)

### BYGRDRPT is did nt repeat, did repeat
fdata$BYGRDRPT<-recode(fdata$BYGRDRPT, "c('1','2','97')='1'")
summary(fdata$BYGRDRPT)

###  BYNONUSG is did not go outside US, did go outside US
fdata$BYNONUSG<-recode(fdata$BYNONUSG, "c('1','2','3','4','5','6','7','8','9','10')='1'")
summary(fdata$BYNONUSG)

### BYSTEXP is now attend 4 year college or less, Bachelors, Masters or higher
fdata$BYSTEXP<-recode(fdata$BYSTEXP, "c('1','2','3','4')='0';c('5')='1';c('6','7')='2'")
summary(fdata$BYSTEXP)

### BYPARASP is now attend 4 year college or less, Bachelors, Masters or higher
fdata$BYPARASP<-recode(fdata$BYPARASP, "c('1','2','3','4')='0';c('5')='1';c('6','7')='2'")
summary(fdata$BYPARASP)

### BYWRKHRS is now no work and work
fdata$BYWRKHRS<-recode(fdata$BYWRKHRS, "c('1','2','3','4','5','6','7','8','9')='1'")
summary(fdata$BYWRKHRS)

### BTEHDEG is now Bachelors or lower, Higher Degree
fdata$BYTEHDEG<-recode(fdata$BYTEHDEG, "c('1','2','3')='1';c('4','5','6','7')='2'")
summary(fdata$BYTEHDEG)

### BTMHDEG is now Bachelors or lower, Higher Degree
fdata$BYTMHDEG<-recode(fdata$BYTMHDEG, "c('1','2','3')='1';c('4','5','6','7')='2'")
summary(fdata$BYTMHDEG)


```

### Setting Up Reference Variable
```{r}
#fdata$BYFCOMP=relevel(factor(fdata$BYFCOMP),ref="1")
#fdata$BYSIBHOM=relevel(factor(fdata$BYSIBHOM),ref="1")
#fdata$BYPARED=relevel(factor(fdata$BYPARED),ref="0")
#fdata$BYINCOME=relevel(factor(fdata$BYINCOME),ref="0")
#fdata$BYGRDRPT=relevel(factor(fdata$BYGRDRPT),ref="0")
#fdata$BYNONUSG=relevel(factor(fdata$BYNONUSG),ref="0")
fdata$BYSTEXP=relevel(factor(fdata$BYSTEXP),ref="0")
fdata$BYPARASP=relevel(factor(fdata$BYPARASP),ref="0")
#fdata$BYWRKHRS=relevel(factor(fdata$BYWRKHRS),ref="0")
#fdata$BYTEHDEG=relevel(factor(fdata$BYTEHDEG),ref="1")
#fdata$BYTMHDEG=relevel(factor(fdata$BYTMHDEG),ref="1")
#fdata$BYREGION=relevel(factor(fdata$BYREGION),ref="3")

```




```{r}
#str(fdata)
fdata[,c(2,3,4,5,6,7,8,11,12,20,21,22,23,24,25,26)]<-sapply(fdata[,c(2,3,4,5,6,7,8,11,12,20,21,22,23,24,25,26)],as.numeric)
str(fdata)
fdata<-na.omit(fdata)
#str(fdata)
```


```{r}
fdata$BYHMWRK<-as.numeric(fdata$BYHMWRK)

mod1<-lm(BYTXMSTD~BYMSEXMTCH+BYMRCEMTCH+BYSTEXP*BYPARASP, data=fdata, na.action=na.omit)
summ.mod1<-summary(mod1)$coefficients[,c(1,3,4)]
write.csv(summ.mod1,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/mod1.csv")
```

```{r}
mod2<-lm(BYTXRSTD~BYESEXMTCH+BYERCEMTCH+BYSTEXP*BYPARASP, data=fdata, na.action=na.omit)
summ.mod2<-summary(mod2)$coefficients[,c(1,3,4)]
#write.csv(summ.mod2,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/mod2.csv")
```

```{r}
mod3<-lm(BYTXMSTD~BYMSEXMTCH+BYMRCEMTCH+BYSTEXP*BYPARASP+BYSTLANG+BYFCOMP+BYSIBHOM+BYPARED+BYINCOME+BYGRDRPT+BYNONUSG+BYSCHPRG+BYTXACC+BYACTCTL+BYSTPREP+BYHMWRK+BYWRKHRS+BYTMHDEG+BYSCTRL+BYURBAN+BYREGION+BYAPIB,data = fdata, na.action = na.omit)
vif(mod3)
summ.mod3<-summary(mod3)$coefficients[,c(1,3,4)]
#write.csv(summ.mod3,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/mod3.csv")
```

```{r}
mod4<-lm(BYTXRSTD~BYESEXMTCH+BYERCEMTCH+BYSTEXP*BYPARASP+BYSTLANG+BYFCOMP+BYSIBHOM+BYPARED+BYINCOME+BYGRDRPT+BYNONUSG+BYSCHPRG+BYTXACC+BYACTCTL+BYSTPREP+BYHMWRK+BYWRKHRS+BYTEHDEG+BYSCTRL+BYURBAN+BYREGION+BYAPIB+BYWRTNGA,data = fdata, na.action = na.omit)
vif(mod4)
summ.mod4<-summary(mod4)$coefficients[,c(1,3,4)]
#write.csv(summ.mod4,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/mod4.csv")
```

```{r}
plot(mod1)
plot(mod2)
plot(mod3)
plot(mod4)
```

```{r}
summary(mod1)
summary(mod2)
summary(mod3)
summary(mod4)
```

```{r}
#confmod1<-confint(mod1)
#write.csv(confmod1,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/conf1mod1.csv")
```

```{r}
#confmod2<-confint(mod2)
#write.csv(confmod2,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/conf1mod2.csv")
```

```{r}
#confmod3<-confint(mod3)
#write.csv(confmod3,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/conf1mod3.csv")
```

```{r}
#confmod4<-confint(mod4)
#write.csv(confmod4,"C:/Users/ssrivas/Desktop/Apurv/UC Davis/UC Davis Courses/Spring Quarter 2021/STA 260/Final Project/conf1mod4.csv")
```

```{r}

levels(fdata$BYPARASP)<-c("High School","Bachelor's","Post-Graduate")
levels(fdata$BYSTEXP)<-c("High School","Bachelor's","Post-Graduate")
interaction.plot(fdata$BYSTEXP,fdata$BYPARASP,fdata$BYTXMSTD, mean, main = "Interaction b/w Parents and Student expectations for Math Scores",xlab = "Students Expectations",ylab = "Math Standardized Scores",trace.label = "Parents Expectations")
interaction.plot(fdata$BYSTEXP,fdata$BYPARASP,fdata$BYTXRSTD, mean, main = "Interaction b/w Parents and Student expectations for Reading Scores",xlab = "Students Expectations",ylab = "Reading Standardized Scores",trace.label = "Parents Exectations")
```

```{r}
nobs(mod3)
```



```{r}
anova(mod1,mod3)
anova(mod2,mod4)
```

```{r BYMSEXMTCH}
par(mfrow=c(1,2))
#math
boxplot(BYTXMSTD~BYMSEXMTCH,data=fdata, main="Plot of Math Score by Sex Match",
   xlab="Sex Match(0=no match,1=match)", ylab="Math Score")

#read
boxplot(BYTXRSTD~BYMSEXMTCH,data=fdata, main="Plot of Reading Score by Sex Match",
   xlab="Sex Match(0=no match,1=match)", ylab="Reading Score")
summary(fdata$BYMSEXMTCH)
fs1=fdata[fdata$BYMSEXMTCH==1,]
summary(fs1$BYTXMSTD)
summary(fs1$BYTXRSTD)
fs0=fdata[fdata$BYMSEXMTCH==0,]
summary(fs0$BYTXMSTD)
summary(fs0$BYTXRSTD)
```
```{r BYERCEMTCH}
par(mfrow=c(1,2))
#math
boxplot(BYTXMSTD~BYERCEMTCH,data=fdata, main="Plot of Math Score by Race Match",
   xlab="Race Match(0=no match,1=match)", ylab="Math Score")

#read
boxplot(BYTXRSTD~BYERCEMTCH,data=fdata, main="Plot of Reading Score by Race Match",
   xlab="Race Match(0=no match,1=match)", ylab="Reading Score")
summary(fdata$BYERCEMTCH)
fr1=fdata[fdata$BYERCEMTCH==1,]
summary(fr1$BYTXMSTD)
summary(fr1$BYTXRSTD)
fr0=fdata[fdata$BYERCEMTCH==0,]
summary(fr0$BYTXMSTD)
summary(fr0$BYTXRSTD)
```
```{r}

levels(fdata$BYERCEMTCH)<-c("mismatch","match")
levels(fdata$BYMSEXMTCH)<-c("mismatch","match")
interaction.plot(fdata$BYERCEMTCH,fdata$BYMSEXMTCH,fdata$BYTXMSTD, mean, main = "Interaction b/w Race and Sex Match for Math Scores",xlab = "Race Match",ylab = "Math Standardized Scores",trace.label = "Sex match")
interaction.plot(fdata$BYERCEMTCH,fdata$BYMSEXMTCH,fdata$BYTXRSTD, mean, main = "Interaction b/w Race and Sex Match for Reading Scores",xlab = "Race Match",ylab = "Reading Standardized Scores",trace.label = "Sex match")
```

```{r BYPARASP}
par(mfrow=c(1,2))
#math
boxplot(BYTXMSTD~BYPARASP,data=fdata, 
   xlab="Parent expectations", ylab="Math Score")

#read
boxplot(BYTXRSTD~BYPARASP,data=fdata,
   xlab="Parent expectations", ylab="Reading Score")
summary(fdata$BYPARASP)
fp0=fdata[fdata$BYPARASP=="High School",]
summary(fp0$BYTXMSTD)
summary(fp0$BYTXRSTD)
fp1=fdata[fdata$BYPARASP=="Bachelor's",]
summary(fp1$BYTXMSTD)
summary(fp1$BYTXRSTD)
fp2=fdata[fdata$BYPARASP=="Post-Graduate",]
summary(fp2$BYTXMSTD)
summary(fp2$BYTXRSTD)
```

```{r BYSTEXP}
par(mfrow=c(1,2))
#math
boxplot(BYTXMSTD~BYSTEXP,data=fdata, 
   xlab="Student expectations", ylab="Math Score")

#read
boxplot(BYTXRSTD~BYSTEXP,data=fdata, 
   xlab="Student expectations", ylab="Reading Score")

summary(fdata$BYSTEXP)
fp0=fdata[fdata$BYSTEXP=="High School",]
summary(fp0$BYTXMSTD)
summary(fp0$BYTXRSTD)
fp1=fdata[fdata$BYSTEXP=="Bachelor's",]
summary(fp1$BYTXMSTD)
summary(fp1$BYTXRSTD)
fp2=fdata[fdata$BYSTEXP=="Post-Graduate",]
summary(fp2$BYTXMSTD)
summary(fp2$BYTXRSTD)

```

```{r means.for.barplots}

read.indian.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 1 & datasub5$BYERCEMTCH == 1)]
read.indian.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 1 & datasub5$BYERCEMTCH == 0)]
read.asian.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 2 & datasub5$BYERCEMTCH == 1)]
read.asian.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 2 & datasub5$BYERCEMTCH == 0)]
read.black.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 3 & datasub5$BYERCEMTCH == 1)]
read.black.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 3 & datasub5$BYERCEMTCH == 0)]
read.hispanicnot.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 4 & datasub5$BYERCEMTCH == 1)]
read.hispanicnot.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 4 & datasub5$BYERCEMTCH == 0)]
read.hispanicspec.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 5 & datasub5$BYERCEMTCH == 1)]
read.hispanicspec.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 5 & datasub5$BYERCEMTCH == 0)]
read.more.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 6 & datasub5$BYERCEMTCH == 1)]
read.more.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 6 & datasub5$BYERCEMTCH == 0)]
read.white.yes <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 7 & datasub5$BYERCEMTCH == 1)]
read.white.no <- datasub5$BYTXRSTD[which(datasub5$BYRACE == 7 & datasub5$BYERCEMTCH == 0)]

math.indian.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 1 & datasub6$BYMRCEMTCH == 1)]
math.indian.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 1 & datasub6$BYMRCEMTCH == 0)]
math.asian.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 2 & datasub6$BYMRCEMTCH == 1)]
math.asian.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 2 & datasub6$BYMRCEMTCH == 0)]
math.black.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 3 & datasub6$BYMRCEMTCH == 1)]
math.black.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 3 & datasub6$BYMRCEMTCH == 0)]
math.hispanicnot.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 4 & datasub6$BYMRCEMTCH == 1)]
math.hispanicnot.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 4 & datasub6$BYMRCEMTCH == 0)]
math.hispanicspec.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 5 & datasub6$BYMRCEMTCH == 1)]
math.hispanicspec.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 5 & datasub6$BYMRCEMTCH == 0)]
math.more.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 6 & datasub6$BYMRCEMTCH == 1)]
math.more.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 6 & datasub6$BYMRCEMTCH == 0)]
math.white.yes <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 7 & datasub6$BYMRCEMTCH == 1)]
math.white.no <- datasub6$BYTXMSTD[which(datasub6$BYRACE == 7 & datasub6$BYMRCEMTCH == 0)]

math.girls.yes <- datasub8$BYTXMSTD[which(datasub8$BYSEX == 2 & datasub8$BYMSEXMTCH == 1)]
math.girls.no <- datasub8$BYTXMSTD[which(datasub8$BYSEX == 2 & datasub8$BYMSEXMTCH == 0)]
math.boys.yes <- datasub8$BYTXMSTD[which(datasub8$BYSEX == 1 & datasub8$BYMSEXMTCH == 1)]
math.boys.no <- datasub8$BYTXMSTD[which(datasub8$BYSEX == 1 & datasub8$BYMSEXMTCH == 0)]


read.girls.yes <- datasub7$BYTXRSTD[which(datasub7$BYSEX == 2 & datasub7$BYESEXMTCH == 1)]
read.girls.no <- datasub7$BYTXRSTD[which(datasub7$BYSEX == 2 & datasub7$BYESEXMTCH == 0)]
read.boys.yes <- datasub7$BYTXRSTD[which(datasub7$BYSEX == 1 & datasub7$BYESEXMTCH == 1)]
read.boys.no <- datasub7$BYTXRSTD[which(datasub7$BYSEX == 1 & datasub7$BYESEXMTCH == 0)]



```
